\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-06-30 coqimpl.tex
Adapt Sealed Family Section

 Family MonotonicF.
  Final Field f : nat -> nat := non_impl.
  Field p : monotonic f := non_impl.
 SealFamily. 
EndFamily.
\end{minted}
  \end{minipage}
\begin{minipage}[t]{0.45\linewidth}
\begin{minted}[fontsize=\footnotesize,escapeinside=@@]{Coq}
Family I2 extends I.
 Override Field a : nat := 1. (* Override *)
 Inherit b.                   (* Inherit  *)

@@

 Inherits Until MonotonicF.
 Override Family MonotonicF.
  Final Field f : nat -> nat := (+ 1).
  Field p : monotonic f := ...
 EndFamily. 
EndFamily.
\end{minted}
  \end{minipage}
  \caption{Example Code for Inheritance and Reasoning}\label{fig:plugin-example3}
\end{figure}

@@

% For example, we may
% define $"add" : \mathbb{Z} \to \mathbb{Z} \to \mathbb{Z}$ as one field
% and intend to prove its commutativity in a following field $"comm"$.
% But to prove so, we need to know more than just a type interface of
% "add" but also its \textbf{concrete fixpoint definition}. Otherwise, in the
% future, once "add" is overridden into a subtraction function then
% "comm" will have problem to be inherited.\YZ{The sentence reads as if you want to enable overriding "add", whereas what you really want is to expose the definition equality about "add".}
% A more concrete example is in \cref{fig:plugin-example1}, consider the
% case we have "Field b : a = 1 := .." instead. Then our
% compilation and Coq's type checking will both fail: in the context of
% "b" (module type "a_6") we only know "a" is of type "nat" but nothing
% more (and this is necessary so that "a" can be overridden in the
% future).
Take the field "c" in "Family I" in \cref{fig:plugin-example3} as the
example. We can see that the first attempt of proving "a = 1" fails.
This should fail, actually---because we want to retain the possibility
to override "a" in the future, but the mere existence of the field "c"
will stop that, since "c" witnesses "a" fix to "1". From the other
perspective, the proof of "c" will require a concrete definition of "a"
instead of just the exposure of its type. \textbf{We can see a conflict
between \textit{overridable} and \textit{(definition-relevant)
reasoning}}.  We enable the exposure of the concrete definition by
providing "Final" keyword upon "a'".

@@

In our Coq plugin, we can achieve such exposure of definitional equality by
exposing the whole definition directly into the compiled module type. Doing so we also prohibit overriding because
subtraction doesn't have the same definition as $add$. We require the
users to decorate field as `Final' so that the plugin will proceed with
this special treatment.
%\YZ{This para is confusing: which kind of equality is used, propositional or judgmental? Where does 'Final' show up in the example?}\EDJreply{I rephrase it and remove the part on talking about meta-theory. Now everything is about judgemental so I don't have to emphasize judgemental. I wanted to make the example small but I guess I can add one `Final' statement into our example. Don't resolve it until I add it into Figure 2.}

\textbf{Special Fields: Overridable/Sealed Family Simulating Sigma Type.} 
Non-overridable fields are still not expressive enough to achieve the full
extent of dependent type programming. There are cases where overriding and
(definition-relevant) reasoning are both needed. For example, we may want a field to be an overridable monotonic function. We cannot split into two fields
\mintinline{Coq}{f : nat → nat} and \mintinline{Coq}{p : monotonic f},
because to provide
"p" we need to have concrete definition of "f" in the family type, which
makes "f" non-overridable at all. One option is to use sigma type
\mintinline{Coq}{{f | monotonic f}}, but that can be cumbersome when we have
a bundle of stuff with multiple refinements.

To resolve this, we provide family sealing mechanism to simulate sigma type---we allow the
users to specify the compiled module type of a given family, and thus we
can have a \textit{non-overridable} (thus locally transparent) field
located in an \textit{overridable} family.
See, for example, the "MonotonicF" in \cref{fig:plugin-example3}.
We simply seal
\mintinline{Coq}{f : nat → nat} and \mintinline{Coq}{p : monotonic f}
together inside an overridable family that does not allow "f"
to be overridden in any of its derived families.
Because "f" is non-overridable inside the family, its definition is
transparent and can be used to prove "p".
However, because the enclosing family of "f" is overridable, to swap the
implementation of "f", we just override the whole ``sealed'' family,
like we do in "I2.MonotonicF". And the compiled module type of
"MonotonicF" will be fixed throughout the extension in different
families---it is always these two fields "f, p".\YZ{Can an overriding, sealed family define new fields not in its parent?}\EDJreply{I don't find this a problem but I didn't support it in my plugin. }

In our example, the seal family "I.MonotonicF" is actually simulating
interfaces as well.  We have
\mbox{``\mintinline{Coq}{Axiom non_impl : forall {T : Type}, T.}''},
and we have assigned it to both "f" and "p", and later instantiate in
"I2" with concrete implementation by overriding this family
"I.MonotonicF".
% Basically, {a : int -> int; property : a is monotonic}, in this case
%   we may want to override "a" using different computation, but we still wnat
%   this refinement/property

>>>>

 Family MonotonicF.
  Field f : nat -> nat := non_impl.
  Field p : monotonic f := non_impl.
 SealFamily. 
EndFamily.
\end{minted}
  \end{minipage}
\begin{minipage}[t]{0.45\linewidth}
\begin{minted}[fontsize=\footnotesize,escapeinside=@@]{Coq}
Family I2 extends I.
 Override Field a : nat := 1. (* Override *)
 Inherit b.                   (* Inherit  *)

@@

 Inherits Until MonotonicF.
 Override Family MonotonicF.
  Field f : nat -> nat := (+ 1).
  Field p : monotonic f := ...
 EndFamily. 
EndFamily.
\end{minted}
  \end{minipage}
  \caption{Example Code for Inheritance and Reasoning}\label{fig:plugin-example3}
\end{figure}

@@

% For example, we may
% define $"add" : \mathbb{Z} \to \mathbb{Z} \to \mathbb{Z}$ as one field
% and intend to prove its commutativity in a following field $"comm"$.
% But to prove so, we need to know more than just a type interface of
% "add" but also its \textbf{concrete fixpoint definition}. Otherwise, in the
% future, once "add" is overridden into a subtraction function then
% "comm" will have problem to be inherited.\YZ{The sentence reads as if you want to enable overriding "add", whereas what you really want is to expose the definition equality about "add".}
% A more concrete example is in \cref{fig:plugin-example1}, consider the
% case we have "Field b : a = 1 := .." instead. Then our
% compilation and Coq's type checking will both fail: in the context of
% "b" (module type "a_6") we only know "a" is of type "nat" but nothing
% more (and this is necessary so that "a" can be overridden in the
% future).
Take the field "c" in "Family I" in \cref{fig:plugin-example3} as the
example. We can see that the first attempt of proving "a = 1" fails.
\textit{This should fail}---because we want to retain the possibility
to override "a" in the future, but the mere existence of the field "c"
will stop that, since "c" witnesses "a" fix to "1". In our implementation, \textit{this will fail}---the proof of "c" will require a concrete definition of "a",
but our compilation will type check \mintinline{Coq}{c : a = 1 := eq_refl} in the context where we only know \mintinline{Coq}{a : nat} thus failed.

To enable the exposure of the concrete definition,
our plugin provide "Final" keyword. Once we decorate it upon "a'", the second attempt \mintinline{Coq}{c : a' = 1 := eq_refl} will succeed. Contrary to "a", our "a'" is not overridable in "Family I2".

@@

We implement "Final" by
exposing the whole definition directly into the compiled module type. Doing so we will prohibit overriding because
subtraction doesn't have the same definition as $add$. We require the
users to manual decorate field as `Final' so that the plugin will proceed with this special treatment.
%\YZ{This para is confusing: which kind of equality is used, propositional or judgmental? Where does 'Final' show up in the example?}\EDJreply{I rephrase it and remove the part on talking about meta-theory. Now everything is about judgemental so I don't have to emphasize judgemental. I wanted to make the example small but I guess I can add one `Final' statement into our example. Don't resolve it until I add it into Figure 2.}

\textbf{Special Fields: Overridable/Sealed Family Simulating Sigma Type.} 
Non-overridable fields are still not expressive enough to achieve the full
extent of dependent type programming. There are cases where overriding and
(definition-relevant) reasoning are both needed. For example, we may want a field to be an overridable monotonic function. We cannot split into two fields
\mintinline{Coq}{f : nat → nat} and \mintinline{Coq}{p : monotonic f},
because to prove
the latter we need to have concrete definition of "f" in the family type, which
makes "f" non-overridable at all. One option is to use sigma type
\mintinline{Coq}{{f | monotonic f}}, but that can be cumbersome when we have
a bundle of stuff with multiple refinements.

To resolve this, we provide \textit{family sealing} mechanism to simulate the sigma type. In a sealed family, every field will expose its definition inside the family but a family as a whole will be sealed into a relaxed type. Thus we can override one sealed family as a whole and thus change the internal implementation.

For example, the "MonotonicF" in \cref{fig:plugin-example3}:
we simply seal
\mintinline{Coq}{f : nat → nat} and \mintinline{Coq}{p : monotonic f}
together inside an sealed family.
Because the definition of "f" is exposed inside the family, 
its definition is
transparent and can be used to prove "p".
However, because the enclosing family of "f" is overridable, to swap the
implementation of "f", we just override the whole ``sealed'' family,
like we do in "I2.MonotonicF".

And the compiled module type of
"MonotonicF" will be fixed throughout the extension in different
families. Overriding field will always have these two fields "f, p".\YZ{Can an overriding, sealed family define new fields not in its parent?}\EDJreply{I don't find this a problem but I didn't support it in my plugin. }

In our example, the seal family "I.MonotonicF" is actually simulating
interfaces as well.  We have
\mbox{``\mintinline{Coq}{Axiom non_impl : forall {T : Type}, T.}''},
and we have assigned it to both "f" and "p", and later instantiate in
"I2" with concrete implementation by overriding this family
"I.MonotonicF".
% Basically, {a : int -> int; property : a is monotonic}, in this case
%   we may want to override "a" using different computation, but we still wnat
%   this refinement/property
The implementation of sealed family is not so different from that of the normal family, but we need to be careful about the compiled module type.
